On particular diameter bounds for integral point sets in higher dimensions \footnote{ This work was carried out at Voronezh State University and supported by Russian Science Foundation grant 19-11-00197. }

### 2025-02-25 AvdMomZvol_rails/Article.tex
[rails] More references

\begin{dfn}
	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
	a \textit{facher} set.
\end{dfn}
Most of the examples of planar integral point sets turns to be facher sets.
In~\cite{antonov2008maximal}, facher sets with characteristic 1 are called \textit{semi-crabs}.
For $9 \leq n \leq 122$, the diameter $d(2,n)$ is attained on a facher point set~\cite{kurz2008minimum}.

@@

Circular sets are very important examples
of integral point sets~\cite{harborth1993upper,piepmeyer1996maximum,bat2018number};
for example, the best known upper bound for $d(2,n)$
is attained on a circular set.

@@

Most of the known planar integral point sets fall into these six classes;
however, there are some sophisticated examples which do not~\cite{kreisel2008heptagon,kurz2013constructing,avdeev2021particular}.
A lower bound for the diameter of a PIPS with no collinear triples (so-called PIPS \textit{in semi-general position})
is presented in~\cite{my-semi-general-5-4-bound-2019}.

@@

There is still no general construction for a rails or scissors PIPS of arbitrary cardinality.
For rails PIPSs, we can conjecture that there exists a set of arbitrary cardinality, with 2 points on one line
and all the rest on the other;
on the other hand, it is still unknown whether there are any rails PIPSs with 4 and 5 points on the lines.

>>>>

\begin{dfn}
	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
	a \textit{facher} set.
\end{dfn}
Most of the examples of planar integral point sets turns to be facher sets.
In~\cite{antonov2008maximal}, facher sets with characteristic 1 are called \textit{semi-crabs}.
For $9 \leq n \leq 122$, the diameter $d(2,n)$ is attained on a facher point set~\cite{kurz2008minimum}.
Facher sets have relatively simple structure and thus
they are rather well-investigated~\cite{zvolinsky2021facher,lushina2021vestnikpmm}.

@@

Circular sets are very important examples
of integral point sets~\cite{harborth1993upper,piepmeyer1996maximum,bat2018number};
for example, the best known upper bound for $d(2,n)$
is attained on a circular set.
Moreover, the first occurrence of integral point sets in modern literature (or at least the first one that the authors are aware of) is Anning's article from 1915 which focuses exactly on circular integral point sets~\cite{anning1915discussions}.
For the most recent advances in circular IPS, we refer the reader to~\cite{undrakh2022refinement}.

@@

Most of the known planar integral point sets fall into these six classes;
however, there are some sophisticated examples which do not~\cite{kreisel2008heptagon,kurz2013constructing,avdeev2021particular}.
A lower bound for the diameter of a PIPS with no collinear triples (so-called PIPS \textit{in semi-general position})
is presented in~\cite{my-semi-general-5-4-bound-2019}.
It is important to note that the cardinality of an IPS in semi-general position is bounded
not only by its diameter (i.e. the maximal distance between its points),
but also by the minimal distance.
Solymosi established such a bound~\cite{solymosi2003note};
in particular, he proved that if an IPS in semi-general position has distance 2,
its cardinality cannot exceed 7.
Recently, this result has been refined to 5 points~\cite{lushina2024cardinality}.

@@

There is still no general construction for a rails or scissors PIPS of arbitrary cardinality.
For rails PIPSs, we can conjecture that there exists a set of arbitrary cardinality, with 2 points on one line
and all the rest on the other;
for an example of cardinality 106, we refer the reader to~\cite{momot2022example}.
On the other hand, it is still unknown whether there are any rails PIPSs with 4 and 5 points on the lines.
